---
title: "16S rRNA Copy Number Analysis for Antibiotics Treated Millipede Microbial Communities"
subtitle: "Copy Number Analysis"
author: "Nweze Julius"
date: "`r Sys.Date()`"
link-citations: yes
csl: fems-microbiology-ecology.csl
output:
  rmarkdown::html_document:
    toc: true
    toc_float: true
    toc_depth: 5
    keep_md: true
    number_sections: false
    highlight: "pygments"
    theme: "flatly"
    dev: "png"
    df_print: "kable"
    fig_caption: true
    code_folding: "show"
editor_options: 
  chunk_output_type: console
---

```{r libraries, include=F}
# Load libraries
#.libPaths(c('~/R/library', .libPaths())) # Uncomment if you have no write access to R path

repo <- "http://cran.wu.ac.at"
lib.loc <- Sys.getenv("R_LIBS_USER")

update.packages(
    lib.loc, 
    repos = repo,
    ask = FALSE
)

.cran_libs <- c(
  "knitr", # A General-Purpose Package for Dynamic Report Generation in R
  "kableExtra", # Construct Complex Table with 'kable' and Pipe Syntax
  "rmarkdown", # Dynamic Documents for R
  "extrafont", # for extra figure fonts
  "tidyverse", # for dplyr forcats ggplot2 readr tibble
  "grid", # The Grid Graphics Package
  "magrittr", # pipes
  "scales", # Generic plot scaling methods
  "svglite", # for svg files
  "vegan",
  "performance",
  "patchwork",
  "car", # Companion to Applied Regression
  "rcompanion", #Functions to Support Extension Education Program Evaluation
  "multcomp", # Simultaneous Inference in General Parametric Models 
  "nlme", # Fit Linear Model Using Generalized Least Squares
  "ggResidpanel", # Panels and Interactive Versions of Diagnostic Plots using 
  "lsmeans", # Least-Squares Means
  "survminer", # For Survival Analysis
  "ranger",
  "lmtest",
  "ggfortify",
  "gtsummary", 
   "cmprsk",  # Competing Risks Packages
  "sm",  # Smooth survival plot - quantile of survival
  "prodlim",
  "pec"
) 

.inst <- .cran_libs %in% installed.packages()
if (any(!.inst)) {
   install.packages(.cran_libs[!.inst],
                    repos = repo,
                    lib = lib.loc)
}

.bioc_libs <- c(
  #"multtest", #Resampling-based multiple hypothesis testing
)

.bioc_inst <- .bioc_libs %in% installed.packages()
if (any(!.bioc_inst)) {
   if (!requireNamespace("BiocManager", quietly = TRUE))
   install.packages("BiocManager")
   BiocManager::install(ask = F, lib = lib.loc)  # upgrade bioC packages
   BiocManager::install(.bioc_libs[!.bioc_inst], ask = F, lib = lib.loc)
}

.local_libs <- c()

.inst <- names(.local_libs) %in% installed.packages()
if (any(!.inst)) {
   install.packages(paste0("~/R/", .local_libs[!.inst]) ,repos = NULL, type = "source", lib = lib.loc)
}

.github_libs <- c()

.github_lib_names <- stringr::str_replace(.github_libs, ".*/(.*)$", "\\1")

.github_inst <- .github_lib_names %in% installed.packages()
if (any(!.github_inst)) {
  devtools::install_github(.github_libs[!.github_inst],
                           lib = lib.loc,
                           dependencies = TRUE)
}

# Load packages into session, and print package version
(loaded.libs <- sapply(c(.cran_libs, .bioc_libs, names(.local_libs), .github_lib_names), require, character.only = TRUE))
if (!all(loaded.libs)) {stop(paste("Package(s):", names(loaded.libs[loaded.libs == FALSE]), "could not be loaded"))}
sapply(c(.cran_libs, .bioc_libs, names(.local_libs), .github_lib_names), packageVersion)
```

```{r style settings, include=F}
options(width = 90, knitr.table.format = "html") 
opts_chunk$set(
  warning = FALSE,
  message = FALSE,
  cache = TRUE,
  dev = "svglite",
  fig.ext = "svg",
  dpi = 300,
#  fig.width = 12,
#  fig.height = 8,
  cache.path = "Copy_number_cache/",
  fig.path = "Copy_number_figs/"
)
f_name <- "DejaVu Sans" #sub("\\s//", "", f_name)
f_size <- 14
font_import(pattern = "DejaVuSans", prompt = FALSE)
loadfonts() # registers fonts
theme_set(theme_bw(base_size = f_size, base_family = f_name)) # set theme for plots
pom4 <- ggpomological:::pomological_palette[c(2, 9, 3, 1)] # set colours
```

```{r functions, include=F}
PlotLmResid <- function(lm.df, which = c(1:6), mfrow = c(3, 2)){

  if (length(levels(as.factor(lm.df$.fitted))) < 10) {# if number of unique x values is <10 just draw a line through the means
    smoother <- stat_summary(fun.y = mean, colour = "red", geom = "line")
  } else smoother <- stat_smooth(method = "loess", geom = "smooth", se = FALSE, colour = "firebrick", size = 1)
  
  # residuals vs fitted
  g1 <- ggplot(lm.df, aes(.fitted, .resid)) +
    geom_point()  +
    smoother + 
    geom_hline(yintercept = 0, linetype = 2, size = .2) +
    scale_x_continuous("Fitted Values") +
    scale_y_continuous("Residual") +
    labs(title = "Residuals vs Fitted")
  
  # normal qq
  a <- quantile(lm.df$.stdresid, c(0.25, 0.75), na.rm = TRUE)
  b <- qnorm(c(0.25, 0.75))
  slope <- diff(a)/diff(b)
  int <- a[1] - slope * b[1]
  g2 <- ggplot(lm.df, aes(sample = .stdresid)) +
    stat_qq() +
    geom_abline(slope = slope, intercept = int, colour = "firebrick", size = 1) +
      scale_x_continuous("Theoretical Quantiles") +
      scale_y_continuous("Standardized Quantiles") +
      labs(title = "Normal Q-Q")
 
  # scale-location
  g3 <- ggplot(lm.df, aes(.fitted, sqrt(abs(.stdresid)))) +
    geom_point() +
    smoother +
    scale_x_continuous("Fitted Values") +
    scale_y_continuous("Root of Standardized Residuals") +
    labs(title = "Scale-Location")
 
  # residuals vs leverage
  g4 <- ggplot(lm.df, aes(factors, .stdresid)) +
    geom_point() +
    smoother +
    geom_hline(yintercept = 0, linetype = 2, size = .2) +
    scale_x_continuous("Factor Level Combinations") +
    scale_y_continuous("Standardized Residuals") +
    labs(title = "Residuals vs Factor Levels")
 
#   # cook's distance
#   g5 <-  ggplot(lm.df, aes(rows, .cooksd, ymin=0, ymax=.cooksd)) +
#     geom_point() + geom_linerange() +
#     scale_x_continuous("Observation Number") +
#     scale_y_continuous("Cook's distance") +
#     labs(title="Cook's Distance")  
  
  # cooksd vs leverage
  g5 <- ggplot(lm.df, aes(factors, .cooksd)) +
    geom_point() +
    smoother +
    scale_x_continuous("Factor Level Combinations") +
    scale_y_continuous("Cook's distance") +
    labs(title = "Cook's dist vs Leverage")
  
  # g6 <- PlotACF(lm.df)
  bw <- diff(range(lm.df$.resid)) / (2 * IQR(lm.df$.resid) / length(lm.df$.resid) ^ (1/3))
  sshist <- function(x){ # optimise bins
  # 2006 Author Hideaki Shimazaki
  # Department of Physics, Kyoto University
  # shimazaki at ton.scphys.kyoto-u.ac.jp
	N <- 2 : 100
	C <- numeric(length(N))
	D <- C

	for (i in 1:length(N)) {
		D[i] <- diff(range(x)) / N[i]

		edges = seq(min(x), max(x), length=N[i])
		hp <- hist(x, breaks = edges, plot=FALSE)
		ki <- hp$counts

		k <- mean(ki)
		v <- sum((ki-k) ^ 2) / N[i]

		C[i] <- (2 * k-v) / D[i] ^ 2	#Cost Function
	}

	idx <- which.min(C)
	optD <- D[idx]

	bins <- seq(min(x), max(x), length=N[idx])
	# h = hist(x, breaks = bins)
	# rug(x)

	return(bins)
  }
  
  bins <- sshist(lm.df$.resid)
  g6 <- ggplot(lm.df, aes(.resid)) + 
    geom_histogram(breaks = bins)
 
  plots <- list(g1, g2, g3, g4, g5, g6)
 
  # making the plots
  grid.newpage()
 
  if (prod(mfrow) > 1) {
    mypos <- expand.grid(1:mfrow[1], 1:mfrow[2])
    mypos <- mypos[with(mypos, order(Var1)), ]
    pushViewport(viewport(layout = grid.layout(mfrow[1], mfrow[2])))
    formatter <- function(.){}
  } else {
    mypos <- data.frame(matrix(1, length(which), 2))
    pushViewport(viewport(layout = grid.layout(1, 1)))
    formatter <- function(.) {
      .dontcare <- readline("Hit <Return> to see next plot: ")
      grid.newpage()
    }
  }
 
  j <- 1
  for (i in which) {
    formatter()
    print(plots[[i]], vp = viewport(layout.pos.row = mypos[j, ][1], layout.pos.col = mypos[j, ][2]))
    j <- j + 1
  }
}

PlotACF <- function(lm.df){
  ## generate ACF plot for lm and lme
  # compute acf without plotting
  acz <- acf(lm.df$.resid, plot = F)
  # convert to data frame
  acd <- data.frame(lag = acz$lag, acf = acz$acf)
  # use data frame for ggplot
  ggplot(acd, aes(lag, acf)) + 
    geom_bar(colour = "black", fill = "black", stat = "identity", position = "dodge", width = 0.01) +
    geom_point(colour = "black") +
    geom_hline(yintercept = c(0.05, -0.05), linetype = "dashed") +
    geom_hline(yintercept = 0)
}
```
[julius.nweze@bc.cas.cz](mailto: julius.nweze@bc.cas.cz)  

## 16S Copy Number Analysis
Analysis of bacterial 16S rRNA copy numbers in response to antibiotics treatment.
Species tested: *E. pulchripes*





## *E. pulchripes* faecal pellet bacterial 16S copy number analysis
```{r load faecal copy number data by Epibolus, cache = T}
read_csv("11_Epibolus_Faecal-Pellet_Copy_Numbers.csv",
         trim_ws = TRUE,
         col_names = TRUE) %>% # read the csv file
  mutate_at(c("Code",
              "Treatment",
              "Replicate",
              "Individual"),
            ~ (factor(.))) %>% # convert to factor
   mutate(Treatment = fct_relevel(Treatment, "Control", "Sterile", "2X Diluted", "Undiluted"))  -> # arrange factor levels 4 plotting
  Epi_Faecal
```

*Barplot of E. pulchripes faecal pellet bacterial total 16S rRNA Copies for all replicates*
```{r plot all replicate faecal copy numbers by Epibolus, cache=TRUE}
# Barplot for all Total 16S rRNA Copies
ggplot(Epi_Faecal, aes(x = Code, y = log(`Total 16S rRNA Copies`), fill = Treatment, group = Code)) + 
  geom_bar(stat = 'identity', position='dodge') +  xlab("Groups") + scale_y_continuous(label=scientific)
```

*Barplot of E. pulchripes faecal bacterial mean copy numbers**
```{r plot mean+SE faecal copy numbers by Epibolus, cache=TRUE}

# Barplot for Total 16S rRNA Copies
 Epi_Faecal %>%  # the name of the new data frame and the data frame to be summarised
  drop_na() %>% # drop rows with missing values
  group_by(Treatment) %>%   # the grouping variable
  summarise(`Mean Total 16S rRNA Copies` = mean((`Total 16S rRNA Copies`)),  # calculates the mean of each group
            StD = sd((`Total 16S rRNA Copies`)), # calculates the standard deviation of each group
            n = n(),  # calculates the sample size per group
            SE = sd((`Total 16S rRNA Copies`))/sqrt(n())) -> # calculates the standard error of each group
 Total_mean

# Create plot
ggplot(Total_mean, aes(x = Treatment, y = `Mean Total 16S rRNA Copies`)) + 
  geom_bar(stat = 'identity', position='dodge', fill = pom4, aes(fill=Treatment)) + ylab(("Mean Total 16S rRNA Copies")) +
  geom_errorbar(aes(ymin = `Mean Total 16S rRNA Copies` - SE, ymax = `Mean Total 16S rRNA Copies` + SE), width=.2,
                 position=position_dodge(.9)) +
  scale_y_continuous(n.breaks = 5, limits = c(0, 6.1e+11), labels = label_log(digits = 2))
  

# data <- data.frame(
#   Treatment = c("Control", "Sterile", "2X-Diluted", "Undiluted"),
#   Counts = c(42766667, 23100000, 17083333, 12850000)
# )
# 
# # Calculate the percentage decrease for each treatment compared to the control
# data$Percentage_Decrease <- ((data$Counts - data$Counts[1]) / data$Counts[1]) * 100
# 
# # Print the results
# print(data)


```


#### Anova tests
**Compare E. pulchripes faecal total 16S rRNA Copies**
```{r faecal copy number aov by Epibolus, cache=T}
# ANOVA for Total 16S rRNA Copies
Epi_Faecal %>% 
  aov(`Total 16S rRNA Copies` ~ Treatment, .) -> # fit an anova model
  aov_total
  
print(Anova(aov_total)) # print results

aov_total_df <- fortify(aov_total) # make a data frame for plotting
factor.combinations <-
  as.numeric(factor(aov_total_df$Treatment),
             levels = unique(as.character(paste(aov_total_df$Treatment)))
  ) # needed for "residuals vs leverage
mod_data_df_total <-
  cbind(aov_total_df,
        rows = 1:nrow(aov_total_df),
        factors = factor.combinations)

PlotLmResid(mod_data_df_total) # plot residuals



# Log transformation
# ANOVA for Total 16S rRNA Copies
Epi_Faecal %>% 
  aov(log(`Total 16S rRNA Copies`) ~ Treatment, .) -> # fit an anova model
  aov_total
  
print(Anova(aov_total)) # print results

aov_total_df <- fortify(aov_total) # make a data frame for plotting
factor.combinations <-
  as.numeric(factor(aov_total_df$Treatment),
             levels = unique(as.character(paste(aov_total_df$Treatment)))
  ) # needed for "residuals vs leverage
mod_data_df_total <-
  cbind(aov_total_df,
        rows = 1:nrow(aov_total_df),
        factors = factor.combinations)

PlotLmResid(mod_data_df_total) # plot residuals

```
The ANOVA model indicates significant differences in the total copy numbers.


 **Building a regression model for total 16S rRNA Copies**
```{r 16s RNA copies aov, cache=T}
# Building a regression model
Epi_Faecal_model1 <- Epi_Faecal %>% 
                     lm(`Total 16S rRNA Copies` ~ Treatment, data = .)

print(Anova(Epi_Faecal_model1, type = "III"))

# checking model assumptions
check_model(Epi_Faecal_model1)

Epi_Faecal_emm_mod1 <- emmeans(Epi_Faecal_model1, as.formula(paste("", "Treatment", sep = " ~ ")), weights = "equal")
      
# Contrasts, pairwise comparisons, tests, and confidence intervals.
Epi_Faecal_emm_mod_cons1 <- contrast(Epi_Faecal_emm_mod1, method = "pairwise", adjust = "BH")
print(Epi_Faecal_emm_mod_cons1)



# The mode shows no statistical significance but before moving on, we need to test the model's assumptions. Obviously the dataset does not fit the assumptions of the model so we cannot use the results and we need to transform the model so that the residuals are normally distributed. One of the easiest ways is to log-transform the data before the analysis.

Epi_Faecal %<>%
  mutate(log_copies = log2(`Total 16S rRNA Copies`))

# check the model
Epi_Faecal_model2 <- Epi_Faecal %>% 
                     lm(log_copies ~ Treatment, data = .)
print(Anova(Epi_Faecal_model2, type = "III"))


# checking model assumptions
check_model(Epi_Faecal_model2)

Epi_Faecal_emm_mod2 <- emmeans(Epi_Faecal_model2, as.formula(paste("", "Treatment", sep = " ~ ")), weights = "equal")
      
# Contrasts, pairwise comparisons, tests, and confidence intervals.
Epi_Faecal_emm_mod_cons2 <- contrast(Epi_Faecal_emm_mod2, method = "pairwise", adjust = "BH")
print(Epi_Faecal_emm_mod_cons2)









# Test for heteroscedasticity, we can perform a Breusch-Pagan test
# perform Breusch-Pagan test
bptest(Epi_Faecal_model1)

# Perform Weighted Least Squares Regression
#define weights to use
wt <- 1 / lm(abs(Epi_Faecal_model1$residuals) ~ Epi_Faecal_model1$fitted.values)$fitted.values^2

#perform weighted least squares regression
wls_model <- lm(`Total 16S rRNA Copies` ~ Treatment, data = Epi_Faecal, weights=wt)

#view summary of model
summary(wls_model)


# For transformed
# Test for heteroscedasticity, we can perform a Breusch-Pagan test
# perform Breusch-Pagan test
bptest(Epi_Faecal_model2)

# Perform Weighted Least Squares Regression
#define weights to use
wt <- 1 / lm(abs(Epi_Faecal_model2$residuals) ~ Epi_Faecal_model2$fitted.values)$fitted.values^2

#perform weighted least squares regression
wls_model <- lm(`Total 16S rRNA Copies` ~ Treatment, data = Epi_Faecal, weights=wt)

#view summary of model
summary(wls_model)
```
