---
title: "Microbiome curing through antibiotics"
subtitle: "Weight loss analysis"
author: "Nweze Julius"
date: "`r Sys.Date()`"
link-citations: yes
csl: fems-microbiology-ecology.csl
output:
  rmarkdown::html_document:
    toc: true
    toc_float: true
    toc_depth: 5
    keep_md: true
    number_sections: false
    highlight: "pygments"
    theme: "flatly"
    dev: "png"
    df_print: "kable"
    fig_caption: true
    code_folding: "show"
---

```{r libraries, include=F}
# Load libraries
#.libPaths(c('~/R/library', .libPaths())) # Uncomment if you have no write access to R path

repo <- "http://cran.wu.ac.at"
lib.loc <- Sys.getenv("R_LIBS_USER")

update.packages(
    lib.loc, 
    repos = repo,
    ask = FALSE
)

.cran_libs <- c(
  "knitr", # A General-Purpose Package for Dynamic Report Generation in R
  "kableExtra", # Construct Complex Table with 'kable' and Pipe Syntax
  "rmarkdown", # Dynamic Documents for R
  "extrafont", # for extra figure fonts
  "tidyverse", # for dplyr forcats ggplot2 readr tibble
  "grid", # The Grid Graphics Package
  "magrittr", # pipes
  "scales", # Generic plot scaling methods
  "svglite", # for svg files
  "vegan",
  "car", # Companion to Applied Regression
  "rcompanion", #Functions to Support Extension Education Program Evaluation
  "multcomp", # Simultaneous Inference in General Parametric Models 
  "nlme", # Fit Linear Model Using Generalized Least Squares
  "ggResidpanel", # Panels and Interactive Versions of Diagnostic Plots using 
  "lsmeans", # Least-Squares Means
  "survminer", # For Survival Analysis
  "ranger",
  "readr",
  "forcats",
  "dplyr",
  "ggfortify",
  "gtsummary", 
   "cmprsk",  # Competing Risks Packages
  "sm",  # Smooth survival plot - quantile of survival
  "prodlim",
  "pec"
) 

.inst <- .cran_libs %in% installed.packages()
if (any(!.inst)) {
   install.packages(.cran_libs[!.inst],
                    repos = repo,
                    lib = lib.loc)
}

.bioc_libs <- c(
  #"multtest", #Resampling-based multiple hypothesis testing
)

.bioc_inst <- .bioc_libs %in% installed.packages()
if (any(!.bioc_inst)) {
   if (!requireNamespace("BiocManager", quietly = TRUE))
   install.packages("BiocManager")
   BiocManager::install(ask = F, lib = lib.loc)  # upgrade bioC packages
   BiocManager::install(.bioc_libs[!.bioc_inst], ask = F, lib = lib.loc)
}

.local_libs <- c()

.inst <- names(.local_libs) %in% installed.packages()
if (any(!.inst)) {
   install.packages(paste0("~/R/", .local_libs[!.inst]) ,repos = NULL, type = "source", lib = lib.loc)
}

.github_libs <- c()

.github_lib_names <- stringr::str_replace(.github_libs, ".*/(.*)$", "\\1")

.github_inst <- .github_lib_names %in% installed.packages()
if (any(!.github_inst)) {
  devtools::install_github(.github_libs[!.github_inst],
                           lib = lib.loc,
                           dependencies = TRUE)
}

# Load packages into session, and print package version
(loaded.libs <- sapply(c(.cran_libs, .bioc_libs, names(.local_libs), .github_lib_names), require, character.only = TRUE))
if (!all(loaded.libs)) {stop(paste("Package(s):", names(loaded.libs[loaded.libs == FALSE]), "could not be loaded"))}
sapply(c(.cran_libs, .bioc_libs, names(.local_libs), .github_lib_names), packageVersion)
```

```{r style settings, include=F}
options(width = 90, knitr.table.format = "html") 
opts_chunk$set(
  warning = FALSE,
  message = FALSE,
  cache = TRUE,
  dev = "svglite",
  fig.ext = "svg",
  dpi = 300,
#  fig.width = 12,
#  fig.height = 8,
  cache.path = "Weight_loss_cache/",
  fig.path = "Weight_loss_figs/"
)
f_name <- "DejaVu Sans" #sub("\\s//", "", f_name)
f_size <- 14
font_import(pattern = "DejaVuSans", prompt = FALSE)
loadfonts() # registers fonts
theme_set(theme_bw(base_size = f_size, base_family = f_name)) # set theme for plots
pom4 <- ggpomological:::pomological_palette[c(2, 9, 3, 1)] # set colours
```

```{r functions, include=F}
PlotLmResid <- function(lm.df, which = c(1:6), mfrow = c(3, 2)){

  if (length(levels(as.factor(lm.df$.fitted))) < 10) {# if number of unique x values is <10 just draw a line through the means
    smoother <- stat_summary(fun.y = mean, colour = "red", geom = "line")
  } else smoother <- stat_smooth(method = "loess", geom = "smooth", se = FALSE, colour = "firebrick", size = 1)
  
  # residuals vs fitted
  g1 <- ggplot(lm.df, aes(.fitted, .resid)) +
    geom_point()  +
    smoother + 
    geom_hline(yintercept = 0, linetype = 2, size = .2) +
    scale_x_continuous("Fitted Values") +
    scale_y_continuous("Residual") +
    labs(title = "Residuals vs Fitted")
  
  # normal qq
  a <- quantile(lm.df$.stdresid, c(0.25, 0.75), na.rm = TRUE)
  b <- qnorm(c(0.25, 0.75))
  slope <- diff(a)/diff(b)
  int <- a[1] - slope * b[1]
  g2 <- ggplot(lm.df, aes(sample = .stdresid)) +
    stat_qq() +
    geom_abline(slope = slope, intercept = int, colour = "firebrick", size = 1) +
      scale_x_continuous("Theoretical Quantiles") +
      scale_y_continuous("Standardized Quantiles") +
      labs(title = "Normal Q-Q")
 
  # scale-location
  g3 <- ggplot(lm.df, aes(.fitted, sqrt(abs(.stdresid)))) +
    geom_point() +
    smoother +
    scale_x_continuous("Fitted Values") +
    scale_y_continuous("Root of Standardized Residuals") +
    labs(title = "Scale-Location")
 
  # residuals vs leverage
  g4 <- ggplot(lm.df, aes(factors, .stdresid)) +
    geom_point() +
    smoother +
    geom_hline(yintercept = 0, linetype = 2, size = .2) +
    scale_x_continuous("Factor Level Combinations") +
    scale_y_continuous("Standardized Residuals") +
    labs(title = "Residuals vs Factor Levels")
 
#   # cook's distance
#   g5 <-  ggplot(lm.df, aes(rows, .cooksd, ymin=0, ymax=.cooksd)) +
#     geom_point() + geom_linerange() +
#     scale_x_continuous("Observation Number") +
#     scale_y_continuous("Cook's distance") +
#     labs(title="Cook's Distance")  
  
  # cooksd vs leverage
  g5 <- ggplot(lm.df, aes(factors, .cooksd)) +
    geom_point() +
    smoother +
    scale_x_continuous("Factor Level Combinations") +
    scale_y_continuous("Cook's distance") +
    labs(title = "Cook's dist vs Leverage")
  
  # g6 <- PlotACF(lm.df)
  bw <- diff(range(lm.df$.resid)) / (2 * IQR(lm.df$.resid) / length(lm.df$.resid) ^ (1/3))
  sshist <- function(x){ # optimise bins
  # 2006 Author Hideaki Shimazaki
  # Department of Physics, Kyoto University
  # shimazaki at ton.scphys.kyoto-u.ac.jp
	N <- 2 : 100
	C <- numeric(length(N))
	D <- C

	for (i in 1:length(N)) {
		D[i] <- diff(range(x)) / N[i]

		edges = seq(min(x), max(x), length=N[i])
		hp <- hist(x, breaks = edges, plot=FALSE)
		ki <- hp$counts

		k <- mean(ki)
		v <- sum((ki-k) ^ 2) / N[i]

		C[i] <- (2 * k-v) / D[i] ^ 2	#Cost Function
	}

	idx <- which.min(C)
	optD <- D[idx]

	bins <- seq(min(x), max(x), length=N[idx])
	# h = hist(x, breaks = bins)
	# rug(x)

	return(bins)
  }
  
  bins <- sshist(lm.df$.resid)
  g6 <- ggplot(lm.df, aes(.resid)) + 
    geom_histogram(breaks = bins)
 
  plots <- list(g1, g2, g3, g4, g5, g6)
 
  # making the plots
  grid.newpage()
 
  if (prod(mfrow) > 1) {
    mypos <- expand.grid(1:mfrow[1], 1:mfrow[2])
    mypos <- mypos[with(mypos, order(Var1)), ]
    pushViewport(viewport(layout = grid.layout(mfrow[1], mfrow[2])))
    formatter <- function(.){}
  } else {
    mypos <- data.frame(matrix(1, length(which), 2))
    pushViewport(viewport(layout = grid.layout(1, 1)))
    formatter <- function(.) {
      .dontcare <- readline("Hit <Return> to see next plot: ")
      grid.newpage()
    }
  }
 
  j <- 1
  for (i in which) {
    formatter()
    print(plots[[i]], vp = viewport(layout.pos.row = mypos[j, ][1], layout.pos.col = mypos[j, ][2]))
    j <- j + 1
  }
}

PlotACF <- function(lm.df){
  ## generate ACF plot for lm and lme
  # compute acf without plotting
  acz <- acf(lm.df$.resid, plot = F)
  # convert to data frame
  acd <- data.frame(lag = acz$lag, acf = acz$acf)
  # use data frame for ggplot
  ggplot(acd, aes(lag, acf)) + 
    geom_bar(colour = "black", fill = "black", stat = "identity", position = "dodge", width = 0.01) +
    geom_point(colour = "black") +
    geom_hline(yintercept = c(0.05, -0.05), linetype = "dashed") +
    geom_hline(yintercept = 0)
}
```
[roey.angel@bc.cas.cz](mailto: roey.angel@bc.cas.cz)  

## Weight loss analysis
Analysis of animal weight loss over time in response to antibiotics treatment.
Species tested: *E. pulchripes* and *G. connexa*.

### *E. pulchripes* analysis 

```{r load E. pulchripes data, cache = T}
Epibolus_weights <- read_csv("1_Epibolus_weight.csv",
                             trim_ws = TRUE,
                             col_names = TRUE) %>% 
  mutate(across(c("Code",
                  "Treatment",
                  "Replicate",
                  "Individual"),
                ~ factor(.))) %>% 
  mutate(across(c("Date"), ~ as.Date(., format = "%d/%m/%Y"))) %>% 
  mutate(Treatment = fct_relevel(Treatment, "Control", "Sterile", "2X Diluted", "Undiluted"))
```

#### Plot time series

**Plot time series of all replicates**
```{r Epibolus plot all weights, cache=TRUE}
ggplot(Epibolus_weights, aes(x = Day, y = `Weight (g)`, colour = Treatment, group = Code)) + 
  geom_point() +
  geom_line() +
  scale_colour_manual(values = pom4, name = "") 
```

**Plot time series of means**
```{r Epibolus plot mean+SE weights, cache=TRUE}
Epibolus_weights %>%  # the name of the new data frame and the data frame to be summarised
  drop_na() %>% # drop rows with missing values
  group_by(Treatment, Day) %>%   # the grouping variable
  summarise(`Mean weight (g)` = mean(`Weight (g)`),  # calculates the mean of each group
            StD = sd(`Weight (g)`), # calculates the standard deviation of each group
            n = n(),  # calculates the sample size per group
            SE = sd(`Weight (g)`)/sqrt(n())) -> # calculates the standard error of each group
  Epibolus_weights_mean

# Create plot
ggplot(Epibolus_weights_mean, aes(x = Day, y = `Mean weight (g)`)) +
  geom_point(aes(colour = Treatment), size = 6) +
  geom_ribbon(aes(
    ymin = `Mean weight (g)` - SE,
    ymax = `Mean weight (g)` + SE,
    fill = Treatment
  ),
  alpha = 1 / 4) +
  geom_line(aes(colour = Treatment)) +
  labs(y = "Weight (g)") +
  scale_colour_manual(values = pom4, name = "") +
  scale_fill_manual(values = pom4, name = "") 
```

```{r Epibolus table mean+SE weights, cache=TRUE}
Epibolus_weights_mean %>% # output a table of group means
  kable(., digits = c(0, 1, 1, 1, 0, 1)) %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive"), full_width = F)
```

#### Anova tests
**Compare time 0 weights between groups**
```{r Epibolus t0 weight aov, cache=T}
Epibolus_weights %>% 
  filter(., Day  == 0) %>% # select only day 0 entries 
  aov(`Weight (g)` ~ Treatment, .) -> # fit an anova model
  aov_t0
  
print(Anova(aov_t0)) # print results

aov_t0_df <- fortify(aov_t0) # make a data frame for plotting
factor.combinations <-
  as.numeric(factor(aov_t0_df$Treatment),
             levels = unique(as.character(paste(aov_t0_df$Treatment)))
  ) # needed for "residuals vs leverage
mod_data_df <-
  cbind(aov_t0_df,
        rows = 1:nrow(aov_t0_df),
        factors = factor.combinations)

PlotLmResid(mod_data_df) # plot residuals
```

The ANOVA model indicates no significant different in the weights of the animals at time 0

**Post-hoc tests for time 0**
```{r Epibolus t0 weight post-hoc, cache=T}
TukeyHSD(aov_t0) # Compute Tukey Honest Significant Differences
```



#### Anova tests
**Compare time 7 weights between groups**
```{r Epibolus t7 weight aov, cache=T}
Epibolus_weights %>% 
  filter(., Day  == 7) %>% # select only day 7 entries 
  aov(`Weight (g)` ~ Treatment, .) -> # fit an anova model
  aov_t7
  
print(Anova(aov_t7)) # print results

aov_t7_df <- fortify(aov_t7) # make a data frame for plotting
factor.combinations <-
  as.numeric(factor(aov_t7_df$Treatment),
             levels = unique(as.character(paste(aov_t7_df$Treatment)))
  ) # needed for "residuals vs leverage
mod_data_df <-
  cbind(aov_t7_df,
        rows = 1:nrow(aov_t7_df),
        factors = factor.combinations)

PlotLmResid(mod_data_df) # plot residuals
```

The ANOVA model indicates no significant different in the weights of the animals at time 7

**Post-hoc tests for time 7**
```{r Epibolus t7 weight post-hoc, cache=T}
TukeyHSD(aov_t7) # Compute Tukey Honest Significant Differences
```
Also in the post-hoc tests no pair of treatment was found to have a significant difference in weights.


#### Anova tests
**Compare time 14 weights between groups**
```{r Epibolus t14 weight aov, cache=T}
Epibolus_weights %>% 
  filter(., Day  == 14) %>% # select only day 14 entries 
  aov(`Weight (g)` ~ Treatment, .) -> # fit an anova model
  aov_t14
  
print(Anova(aov_t14)) # print results

aov_t14_df <- fortify(aov_t14) # make a data frame for plotting
factor.combinations <-
  as.numeric(factor(aov_t14_df$Treatment),
             levels = unique(as.character(paste(aov_t14_df$Treatment)))
  ) # needed for "residuals vs leverage
mod_data_df <-
  cbind(aov_t14_df,
        rows = 1:nrow(aov_t14_df),
        factors = factor.combinations)

PlotLmResid(mod_data_df) # plot residuals
```

The ANOVA model indicates no significant different in the weights of the animals at time 14

**Post-hoc tests for time 14**
```{r Epibolus t14 weight post-hoc, cache=T}
TukeyHSD(aov_t14) # Compute Tukey Honest Significant Differences
```
Also in the post-hoc tests no pair of treatment was found to have a significant difference in weights.


#### Anova tests
**Compare time 21 weights between groups**
```{r Epibolus t21 weight aov, cache=T}
Epibolus_weights %>% 
  filter(., Day  == 21) %>% # select only day 21 entries 
  aov(`Weight (g)` ~ Treatment, .) -> # fit an anova model
  aov_t21
  
print(Anova(aov_t21)) # print results

aov_t21_df <- fortify(aov_t21) # make a data frame for plotting
factor.combinations <-
  as.numeric(factor(aov_t21_df$Treatment),
             levels = unique(as.character(paste(aov_t21_df$Treatment)))
  ) # needed for "residuals vs leverage
mod_data_df <-
  cbind(aov_t21_df,
        rows = 1:nrow(aov_t21_df),
        factors = factor.combinations)

PlotLmResid(mod_data_df) # plot residuals
```

The ANOVA model indicates no significant different in the weights of the animals at time 21

**Post-hoc tests for time 21**
```{r Epibolus t21 weight post-hoc, cache=T}
TukeyHSD(aov_t21) # Compute Tukey Honest Significant Differences
```
Also in the post-hoc tests no pair of treatment was found to have a significant difference in weights.


#### Anova tests
**Compare time 28 weights between groups**
```{r Epibolus t28 weight aov, cache=T}
Epibolus_weights %>% 
  filter(., Day  == 28) %>% # select only day 28 entries 
  aov(`Weight (g)` ~ Treatment, .) -> # fit an anova model
  aov_t28
  
print(Anova(aov_t28)) # print results

aov_t28_df <- fortify(aov_t28) # make a data frame for plotting
factor.combinations <-
  as.numeric(factor(aov_t28_df$Treatment),
             levels = unique(as.character(paste(aov_t28_df$Treatment)))
  ) # needed for "residuals vs leverage
mod_data_df <-
  cbind(aov_t28_df,
        rows = 1:nrow(aov_t28_df),
        factors = factor.combinations)

PlotLmResid(mod_data_df) # plot residuals
```

The ANOVA model indicates no significant different in the weights of the animals at time 28


**Post-hoc tests for time 28**
```{r Epibolus t28 weight post-hoc, cache=T}
TukeyHSD(aov_t28) # Compute Tukey Honest Significant Differences
```
Also in the post-hoc tests no pair of treatment was found to have a significant difference in weights.



#### Anova tests
**Compare time 35 weights between groups**
```{r Epibolus t35 weight aov, cache=T}
Epibolus_weights %>% 
  filter(., Day  == 35) %>% # select only day 35 entries 
  aov(`Weight (g)` ~ Treatment, .) -> # fit an anova model
  aov_t35
  
print(Anova(aov_t35)) # print results

aov_t35_df <- fortify(aov_t35) # make a data frame for plotting
factor.combinations <-
  as.numeric(factor(aov_t35_df$Treatment),
             levels = unique(as.character(paste(aov_t35_df$Treatment)))
  ) # needed for "residuals vs leverage
mod_data_df <-
  cbind(aov_t35_df,
        rows = 1:nrow(aov_t35_df),
        factors = factor.combinations)

PlotLmResid(mod_data_df) # plot residuals
```

The ANOVA model indicates no significant different in the weights of the animals at time 35

**Post-hoc tests for time 35**
```{r Epibolus t35 weight post-hoc, cache=T}
TukeyHSD(aov_t35) # Compute Tukey Honest Significant Differences
```
Also in the post-hoc tests no pair of treatment was found to have a significant difference in weights.



**Compare time 42 weights between groups**
To do that I'll first subtract the weight at time 42 from time 0 for each individual
```{r Epibolus t42 weight aov, cache=T}
Epibolus_weights %>% 
  filter(., Day  == 0 | Day == 42) %>% # select only day 0 and 42 entries 
  pivot_wider(id_cols = c(Code, Treatment, Replicate, Individual), names_from = Day, values_from = `Weight (g)`) %>% # make a wide format table
  mutate(`Weight diff. (g)` = `42` - `0`) %>% # subtract day 0 from day 42 weight
  aov(`Weight diff. (g)` ~ Treatment, .) -> # fit anova
  aov_t42
  
print(Anova(aov_t42)) # print results

aov_t42_df <- fortify(aov_t42) # make a data frame for plotting

factor.combinations <-
  as.numeric(factor(aov_t42_df$Treatment),
             levels = unique(as.character(paste(aov_t42_df$Treatment)))
  ) # needed for "residuals vs leverage
mod_data_df <-
  cbind(aov_t42_df,
        rows = 1:nrow(aov_t42_df),
        factors = factor.combinations)

PlotLmResid(mod_data_df) # plot residuals
```

The analysis indicates a significant difference in weights at day 42

**Post-hoc tests for time 42**
```{r Epibolus t42 weight post-hoc, cache=T}
TukeyHSD(aov_t42)
```

Pairs: Sterile-Control, 2X diluted-Control and Undiluted-Control have significant differences between them but there's no difference between antibiotic treatment and sterile food!

#### Time-series analysis
So far we've only taken into account the last time point, however we can analyse the entire time series to see if there is a difference in trends.

**Fit a linear mixed-effects model**
```{r Epibolus time-series analysis}
Epibolus_weights %>% 
  drop_na() %>% # drop missing entries
  rename(Weight = `Weight (g)`) -> # rename column 
  Epibolus_weights4model # data frame for modeling

model.a <- 
  lme(Weight ~ Treatment + Day + Treatment * Day, 
      random = ~ 1 | Individual,
      data = Epibolus_weights4model) # determine autocorrelations/lag

autocorr.lme <- ACF(model.a,
                form = ~ Day | Individual)[2, 2]

# Fit a linear mixed-effects model 
lme(Weight ~ Treatment + Day + Treatment * Day, 
    data = Epibolus_weights4model,
    random = ~1|Individual,
    correlation = corAR1(form = ~ Day | Individual, 
                         value = autocorr.lme),
    method = "REML") ->
lme.fit

summary(lme.fit) %>%
   
anova(lme.fit)
```
The model indicates some minor effect of the treatment on certain days


**Plot diagnostic plots for the lme model**
```{r Epibolus diag plots lme, cache=T}
resid_panel(lme.fit)
```

Check if including random effects term is justified
```{r Epibolus rand, cache=T}
# is the random effect meaningful?
Epibolus_weights4model %>% 
  gls(Weight ~ Treatment + Day + Treatment * Day,
      data = .,
      method = "REML") ->
  gls.fixed.fit

anova(lme.fit, gls.fixed.fit)
```
Yes

*Model significance*
Calculate a p-value and pseudo R-squared value for the model by comparing it to a null model with no fixed effects
```{r Epibolus model p-value and post-hoc tests}
Epibolus_weights4model %>% 
  lme(Weight ~ 1,
      random = ~1|Individual,
      data = .) ->
  model.null

nagelkerke(lme.fit,
           model.null)

marginal <- lsmeans(lme.fit, ~Treatment:Day)

cld(marginal,
    alpha   = 0.05,
    Letters = letters,     ### Use lower-case letters for .group
    adjust  = "tukey") %>%     ###  Tukey-adjusted comparisons
  kable(., digits = c(0, 1, 2, 1, 1, 1, 1, 0)) %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive"), full_width = F)
```
