---
title: "mcrA copies from faeces after methanogenesis supression"
author: "Nweze Julius"
date: "`r Sys.Date()`"
output:
  html_document:
    df_print: paged
    toc: yes
    toc_depth: '5'
  rmarkdown::html_document:
    code_folding: show
    dev: png
    df_print: kable
    fig_caption: yes
    highlight: pygments
    keep_md: yes
    number_sections: no
    theme: flatly
    toc: yes
    toc_depth: 5
    toc_float: yes
link-citations: yes
csl: fems-microbiology-ecology.csl
subtitle: mcrA inhibition analysis
editor_options: 
  chunk_output_type: console
---

```{r libraries, include=F}
# Load libraries
#.libPaths(c('~/R/library', .libPaths())) # Uncomment if you have no write access to R path

repo <- "http://cran.wu.ac.at"
lib.loc <- Sys.getenv("R_LIBS_USER")

update.packages(
    lib.loc, 
    repos = repo,
    ask = FALSE
)

.cran_libs <- c(
 "knitr", # A General-Purpose Package for Dynamic Report Generation in R
 "kableExtra", # Construct Complex Table with 'kable' and Pipe Syntax
#  "rmarkdown", # Dynamic Documents for R
  "extrafont", # for extra figure fonts
  "tidyverse", # for dplyr forcats ggplot2 readr tibble
  "readODS", # #Read ODS Files
 "grid", # The Grid Graphics Package
#  "magrittr", # pipes
  "scales", # Generic plot scaling methods
  "svglite", # for svg files
#  "vagen",
  "ggResidpanel",
 "car", # Companion to Applied Regression
 "rcompanion", #Functions to Support Extension Education Program Evaluation
# "multcomp", # Simultaneous Inference in General Parametric Models
 "nlme", # Fit Linear Model Using Generalized Least Squares
 "emmeans", # Estimated Marginal Means, aka Least-Squares Means
"performance" # Assessment of Regression Models Performance 
) 

.inst <- .cran_libs %in% installed.packages()
if (any(!.inst)) {
   install.packages(.cran_libs[!.inst],
                    repos = repo,
                    lib = lib.loc)
}

.bioc_libs <- c(
  #"multtest", #Resampling-based multiple hypothesis testing
)

.bioc_inst <- .bioc_libs %in% installed.packages()
if (any(!.bioc_inst)) {
   if (!requireNamespace("BiocManager", quietly = TRUE))
   install.packages("BiocManager")
   BiocManager::install(ask = F, lib = lib.loc)  # upgrade bioC packages
   BiocManager::install(.bioc_libs[!.bioc_inst], ask = F, lib = lib.loc)
}

.local_libs <- c()

.inst <- names(.local_libs) %in% installed.packages()
if (any(!.inst)) {
   install.packages(paste0("~/R/", .local_libs[!.inst]) ,repos = NULL, type = "source", lib = lib.loc)
}

.github_libs <- c(
  "wilkelab/ggtext", # Improved text rendering support for 'ggplot2' 
  "ACCLAB/dabestr" # Data Analysis using Bootstrap-Coupled Estimation 
)

.github_lib_names <- stringr::str_replace(.github_libs, ".*/(.*)$", "\\1")

.github_inst <- .github_lib_names %in% installed.packages()
if (any(!.github_inst)) {
  devtools::install_github(.github_libs[!.github_inst],
                           lib = lib.loc,
                           dependencies = TRUE)
}

# Load packages into session, and print package version
(loaded.libs <- sapply(c(.cran_libs, .bioc_libs, names(.local_libs), .github_lib_names), require, character.only = TRUE))
if (!all(loaded.libs)) {stop(paste("Package(s):", names(loaded.libs[loaded.libs == FALSE]), "could not be loaded"))}
sapply(c(.cran_libs, .bioc_libs, names(.local_libs), .github_lib_names), packageVersion)
```

```{r style settings, include=F}
options(width = 90, knitr.table.format = "html") 
opts_chunk$set(
  warning = FALSE,
  message = FALSE,
  cache = TRUE,
  dev = "svglite",
  fig.ext = "png",
  dpi = 300,
#  fig.width = 12,
#  fig.height = 8,
  cache.path = "Methanogenesis_cache/",
  fig.path = "Methanogenesis_figs/"
)
f_name <- "DejaVu Sans" #sub("\\s//", "", f_name)
f_size <- 14
font_import(pattern = "DejaVuSans", prompt = FALSE)
loadfonts() # registers fonts
theme_set(theme_bw(base_size = f_size, base_family = f_name)) # set theme for plots
pom4 <- ggpomological:::pomological_palette[c(2, 9, 3, 1)] # set colours
```

```{r functions, include=F}
PlotLmResid <- function(lm.df, which = c(1:6), mfrow = c(3, 2)){

  if (length(levels(as.factor(lm.df$.fitted))) < 10) {# if number of unique x values is <10 just draw a line through the means
    smoother <- stat_summary(fun = mean, colour = "red", geom = "line")
  } else smoother <- stat_smooth(method = "loess", geom = "smooth", se = FALSE, colour = "firebrick", size = 1)
  
  # residuals vs fitted
  g1 <- ggplot(lm.df, aes(.fitted, .resid)) +
    geom_point()  +
    smoother + 
    geom_hline(yintercept = 0, linetype = 2, size = .2) +
    scale_x_continuous("Fitted Values") +
    scale_y_continuous("Residual") +
    labs(title = "Residuals vs Fitted")
  
  # normal qq
  a <- quantile(lm.df$.stdresid, c(0.25, 0.75), na.rm = TRUE)
  b <- qnorm(c(0.25, 0.75))
  slope <- diff(a)/diff(b)
  int <- a[1] - slope * b[1]
  g2 <- ggplot(lm.df, aes(sample = .stdresid)) +
    stat_qq() +
    geom_abline(slope = slope, intercept = int, colour = "firebrick", size = 1) +
      scale_x_continuous("Theoretical Quantiles") +
      scale_y_continuous("Standardized Quantiles") +
      labs(title = "Normal Q-Q")
 
  # scale-location
  g3 <- ggplot(lm.df, aes(.fitted, sqrt(abs(.stdresid)))) +
    geom_point() +
    smoother +
    scale_x_continuous("Fitted Values") +
    scale_y_continuous("Root of Standardized Residuals") +
    labs(title = "Scale-Location")
 
  # residuals vs leverage
  g4 <- ggplot(lm.df, aes(factors, .stdresid)) +
    geom_point() +
    smoother +
    geom_hline(yintercept = 0, linetype = 2, size = .2) +
    scale_x_continuous("Factor Level Combinations") +
    scale_y_continuous("Standardized Residuals") +
    labs(title = "Residuals vs Factor Levels")
 
#   # cook's distance
#   g5 <-  ggplot(lm.df, aes(rows, .cooksd, ymin=0, ymax=.cooksd)) +
#     geom_point() + geom_linerange() +
#     scale_x_continuous("Observation Number") +
#     scale_y_continuous("Cook's distance") +
#     labs(title="Cook's Distance")  
  
  # cooksd vs leverage
  g5 <- ggplot(lm.df, aes(factors, .cooksd)) +
    geom_point() +
    smoother +
    scale_x_continuous("Factor Level Combinations") +
    scale_y_continuous("Cook's distance") +
    labs(title = "Cook's dist vs Leverage")
  
  # g6 <- PlotACF(lm.df)
  bw <- diff(range(lm.df$.resid)) / (2 * IQR(lm.df$.resid) / length(lm.df$.resid) ^ (1/3))
  sshist <- function(x){ # optimise bins
  # 2006 Author Hideaki Shimazaki
  # Department of Physics, Kyoto University
  # shimazaki at ton.scphys.kyoto-u.ac.jp
	N <- 2 : 100
	C <- numeric(length(N))
	D <- C

	for (i in 1:length(N)) {
		D[i] <- diff(range(x)) / N[i]

		edges = seq(min(x), max(x), length=N[i])
		hp <- hist(x, breaks = edges, plot=FALSE)
		ki <- hp$counts

		k <- mean(ki)
		v <- sum((ki-k) ^ 2) / N[i]

		C[i] <- (2 * k-v) / D[i] ^ 2	#Cost Function
	}

	idx <- which.min(C)
	optD <- D[idx]

	bins <- seq(min(x), max(x), length=N[idx])
	# h = hist(x, breaks = bins)
	# rug(x)

	return(bins)
  }
  
  bins <- sshist(lm.df$.resid)
  g6 <- ggplot(lm.df, aes(.resid)) + 
    geom_histogram(breaks = bins)
 
  plots <- list(g1, g2, g3, g4, g5, g6)
 
  # making the plots
  grid.newpage()
 
  if (prod(mfrow) > 1) {
    mypos <- expand.grid(1:mfrow[1], 1:mfrow[2])
    mypos <- mypos[with(mypos, order(Var1)), ]
    pushViewport(viewport(layout = grid.layout(mfrow[1], mfrow[2])))
    formatter <- function(.){}
  } else {
    mypos <- data.frame(matrix(1, length(which), 2))
    pushViewport(viewport(layout = grid.layout(1, 1)))
    formatter <- function(.) {
      .dontcare <- readline("Hit <Return> to see next plot: ")
      grid.newpage()
    }
  }
 
  j <- 1
  for (i in which) {
    formatter()
    print(plots[[i]], vp = viewport(layout.pos.row = mypos[j, ][1], layout.pos.col = mypos[j, ][2]))
    j <- j + 1
  }
}

PlotACF <- function(lm.df){
  ## generate ACF plot for lm and lme
  # compute acf without plotting
  acz <- acf(lm.df$.resid, plot = F)
  # convert to data frame
  acd <- data.frame(lag = acz$lag, acf = acz$acf)
  # use data frame for ggplot
  ggplot(acd, aes(lag, acf)) + 
    geom_bar(colour = "black", fill = "black", stat = "identity", position = "dodge", width = 0.01) +
    geom_point(colour = "black") +
    geom_hline(yintercept = c(0.05, -0.05), linetype = "dashed") +
    geom_hline(yintercept = 0)
}
```
[roey.angel@bc.cas.cz](mailto: roey.angel@bc.cas.cz)  

## Methanogenesis inhibition analysis
Analysis of millipede methane production over time in response to inhibition using BES.
Species tested: *E. pulchripes*.

### *E. pulchripes* analysis 

**Load data **
```{r load E. pulchripes data, cache = T}
read_ods("McrA.ods",
         sheet = "Sheet1") %>% # read the ODS file
  mutate_at(c("Code",
              "Treatment",
              "Replicate",
              "Individual"),
            ~ (factor(.))) %>% # convert to factor
  # mutate_at(c("Date"), ~ as.Date(., "%d/%m/%Y")) %>% # convert to date 
  mutate(Treatment = fct_relevel(Treatment, "Control", "5mM-Na2BES", "10mM-Na2BES")) -> # arrange factor levels 4 plotting
  Epibolus_mcrA
```

#### Plot time series

**Plot time series of all replicates**
```{r plot all mcrA, cache=TRUE}
# Plot day vs mcrA
ggplot(Epibolus_mcrA, aes(x = Code, y = `mcrA (copies uL-1)`, fill = Treatment, group = Code)) + 
  geom_bar(stat = 'identity', position='dodge') +  xlab("Groups") +
  scale_colour_manual(values = pom4, name = "") +  scale_fill_manual(values = pom4, name = "") +
  labs(y = "mcrA (Copy number uL<sup>-1</sup>)")  +
  theme(axis.title.y = element_markdown())
  
```

**Plot time series of means**
```{r plot mean+SE mcrA, cache=TRUE}

library(conflicted)  

library(tidyverse)
conflict_prefer("filter", "dplyr")
conflict_prefer("lag", "dplyr")


Epibolus_mcrA %>%  # the name of the new data frame and the data frame to be summarised
  drop_na() %>% # drop rows with missing values
  group_by(Treatment) %>%   # the grouping variable
  summarise(`Mean mcrA (copies uL-1)` = mean(`mcrA (copies uL-1)`),  # calculates the mean of each group
            StD = sd(`mcrA (copies uL-1)`), # calculates the standard deviation of each group
            n = n(),  # calculates the sample size per group
            SE = sd(`mcrA (copies uL-1)`)/sqrt(n())) -> # calculates the standard error of each group
  Epibolus_mcrA_Epibolus_mean


# Create plot
ggplot(Epibolus_mcrA_Epibolus_mean, aes(x = Treatment, y = `Mean mcrA (copies uL-1)`, fill = Treatment)) + 
  geom_bar(stat = 'identity', position='dodge') +
  geom_errorbar(aes(ymin = `Mean mcrA (copies uL-1)` - SE, ymax = `Mean mcrA (copies uL-1)` + SE), width=.2,
                 position=position_dodge(.9)) +  scale_colour_manual(values = pom4, name = "") +  scale_fill_manual(values = pom4, name = "") + labs(y = "mcrA (copy number uL<sup>-1</sup>)")  +
  theme(axis.title.y = element_markdown()) +
  theme(axis.title.y = element_markdown(), 
        text = element_text(size = 20),  
axis.text = element_text(size = 20))

ggsave("BES_methanogen_mcrA.svg",width = 30, height = 20, units = "cm")


Epibolus_mcrA %>%
lm(`mcrA (copies uL-1)` ~ Treatment, .) -> # fit an anova model
  aov_mcrA

print(Anova(aov_mcrA))

TukeyHSD(aov(aov_mcrA)) # Compute Tukey Honest Significant Differences
```

```{r table mean+SE mcrA, cache=TRUE}
# For mcrA
Epibolus_mcrA_Epibolus_mean %>% # output a table of group means
  kable(., digits = c(0, 1, 1, 1, 0, 1)) %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive"), full_width = F)
```

#### Anova tests
**Compare mcrA between groups**
```{r t0 mcrA aov, cache=T}
Epibolus_mcrA %>% 
  lm(`mcrA (copies uL-1)` ~ Treatment, .) -> # fit an anova model
  aov_t0
  
print(Anova(aov_t0)) # print results
# model performence
r2(aov_t0) # R2
check_model(aov_t0)
model_performance(aov_t0)

aov_t0_df <- fortify(aov_t0) # make a data frame for plotting
factor.combinations <-
  as.numeric(factor(aov_t0_df$Treatment),
             levels = unique(as.character(paste(aov_t0_df$Treatment)))
  ) # needed for "residuals vs leverage
mod_data_df <-
  cbind(aov_t0_df,
        rows = 1:nrow(aov_t0_df),
        factors = factor.combinations)

PlotLmResid(mod_data_df) # plot residuals
```

The ANOVA model indicates a significant difference.

**Post-hoc tests for day 0**
```{r t0 mcrA post-hoc, cache=T}
TukeyHSD(aov(aov_t0)) # Compute Tukey Honest Significant Differences
```

Also in the post-hoc tests, there is a significant difference only in the 10mM-Na2BES and Control.






# Symbiotic ciliate counts
```{r load E. pulchripes ciliate data, cache = T}
read_ods("McrA.ods",
         sheet = "Sheet2") %>% # read the ODS file
  mutate_at(c("Code",
              "Treatment",
              "Replicate"),
            ~ (factor(.))) %>% # convert to factor
  # mutate_at(c("Date"), ~ as.Date(., "%d/%m/%Y")) %>% # convert to date 
  mutate(Treatment = fct_relevel(Treatment, "C-Na-BES", "5mM-Na-BES", "10mM-Na-BES")) -> # arrange factor levels 4 plotting
  Epibolus_ciliate
```

**Plot means ciliate counts**
```{r plot mean+SE cilate, cache=TRUE}
scientific_10 <- function(x,suppress_ones=TRUE) {
    s <- scales::scientific_format()(x)
    ## substitute for exact zeros
    s[s=="0e+00"] <- "0"
    ## regex: [+]?  = "zero or one occurrences of '+'"
    s2 <- gsub("e[+]?", " %*% 10^", s )
    ## suppress 1 x
    if (suppress_ones) s2 <- gsub("1 %\\*% +","",s2)
    parse(text=s2)
}


 Epibolus_ciliate %>%  # the name of the new data frame and the data frame to be summarised
  drop_na() %>% # drop rows with missing values
  group_by(Treatment) %>%   # the grouping variable
  summarise(`Mean ciliate counts (cells mL-1)` = mean(`Ciliate counts (cells mL-1)`),  # calculates the mean of each group
            StD = sd(`Ciliate counts (cells mL-1)`), # calculates the standard deviation of each group
            n = n(),  # calculates the sample size per group
            SE = sd(`Ciliate counts (cells mL-1)`)/sqrt(n())) -> # calculates the standard error of each group
  Epibolus_ciliate_mean


# Create plot
ggplot(Epibolus_ciliate_mean, aes(x = Treatment, y = `Mean ciliate counts (cells mL-1)`, fill = Treatment)) + 
  geom_bar(stat = 'identity', position='dodge') +
  geom_errorbar(aes(ymin = `Mean ciliate counts (cells mL-1)` - SE, ymax = `Mean ciliate counts (cells mL-1)` + SE), width=.2,
                 position=position_dodge(.9)) +  scale_colour_manual(values = pom4, name = "") +  scale_fill_manual(values = pom4, name = "") + labs(y = "Mean ciliate counts (cells mL<sup>-1</sup>)")  +
  theme(axis.title.y = element_markdown()) + theme(legend.position="none") +
  theme(axis.title.y = element_markdown(), 
        text = element_text(size = 20),  
axis.text = element_text(size = 20)) +
scale_y_continuous(label=scientific_10) 
ggsave("BES_Ciliate_counts.svg",width = 30, height = 20, units = "cm")



Epibolus_ciliate_mean %>% # output a table of group means
  kable(., digits = c(0, 1, 1, 1, 0, 1)) %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive"), full_width = F)

# Shapiro-Wilk normality test for C-Na-BES Ciliate counts
Epibolus_ciliate %>% 
with(shapiro.test(`Ciliate counts (cells mL-1)`[Treatment == "C-Na-BES"]))# p = 0.3673

# Shapiro-Wilk normality test for 5mM-Na-BES Ciliate counts
Epibolus_ciliate %>% 
with(shapiro.test(`Ciliate counts (cells mL-1)`[Treatment == "5mM-Na-BES"]))# p = 0.5623

# Shapiro-Wilk normality test for 10mM-Na-BES Ciliate counts
Epibolus_ciliate %>% 
with(shapiro.test(`Ciliate counts (cells mL-1)`[Treatment == "10mM-Na-BES"]))# p = 0.9987

# From the output, two p-values are greater than the significance level 0.05 implying that the distribution of the data are not significantly different from the normal distribution. In other words, we can assume the normality.

# We’ll use F-test to test for homogeneity in variances. This can be performed with the function var.test() as follow:

res.ftest <- var.test(`Ciliate counts (cells mL-1)` ~ Epibolus_ciliate$, data = Epibolus_ciliate)
res.ftest

```




# Endosymbiotic methanogen counts
```{r load E. pulchripes methanogens data, cache = T}
read_ods("McrA.ods",
         sheet = "Sheet3") %>% # read the ODS file
  mutate_at(c("Methanogens",
              "Treatment"),
            ~ (factor(.))) %>% # convert to factor
  # mutate_at(c("Date"), ~ as.Date(., "%d/%m/%Y")) %>% # convert to date 
  mutate(Treatment = fct_relevel(Treatment, "C-Na-BES", "5mM-Na-BES", "10mM-Na-BES")) -> # arrange factor levels 4 plotting
  Epibolus_methanogens
```

**Plot means ciliate counts**
```{r plot mean+SE methanogens, cache=TRUE}
scientific_10 <- function(x,suppress_ones=TRUE) {
    s <- scales::scientific_format()(x)
    ## substitute for exact zeros
    s[s=="0e+00"] <- "0"
    ## regex: [+]?  = "zero or one occurrences of '+'"
    s2 <- gsub("e[+]?", " %*% 10^", s )
    ## suppress 1 x
    if (suppress_ones) s2 <- gsub("1 %\\*% +","",s2)
    parse(text=s2)
}


 Epibolus_methanogens %>%  # the name of the new data frame and the data frame to be summarised
  drop_na() %>% # drop rows with missing values
  group_by(Treatment, Methanogens) %>%   # the grouping variable
  summarise(`Mean methanogen counts` = mean(Counts),  # calculates the mean of each group
            StD = sd(Counts), # calculates the standard deviation of each group
            n = n(),  # calculates the sample size per group
            SE = sd(Counts)/sqrt(n())) -> # calculates the standard error of each group
  Epibolus_methanogens_mean


# Create plot
ggplot( Epibolus_methanogens_mean, aes(x = Treatment, y = `Mean methanogen counts`, fill = Methanogens)) + 
  geom_bar(stat = 'identity', position='dodge') +
  geom_errorbar(aes(ymin = `Mean methanogen counts` - SE, ymax = `Mean methanogen counts` + SE), width=.2,
                 position=position_dodge(.9)) +  scale_colour_manual(values = pom4, name = "") +  scale_fill_manual(values = pom4, name = "") + labs(y = "Mean methanogen counts")  +
  theme(axis.title.y = element_markdown()) + theme(legend.position="top") +
  theme(axis.title.y = element_markdown(), 
        text = element_text(size = 20),  
axis.text = element_text(size = 20)) 
ggsave("BES_Methanogen_counts.svg",width = 30, height = 20, units = "cm")
```