---
title: "Microbiome curing through antibiotics"
subtitle: "Survival analysis analysis"
author: "Nweze Julius"
date: "`r Sys.Date()`"
link-citations: yes
csl: fems-microbiology-ecology.csl
output:
  rmarkdown::html_document:
    toc: true
    toc_float: true
    toc_depth: 5
    keep_md: true
    number_sections: false
    highlight: "pygments"
    theme: "flatly"
    dev: "png"
    df_print: "kable"
    fig_caption: true
    code_folding: "show"
---

```{r libraries, include=F}
# Load libraries
#.libPaths(c('~/R/library', .libPaths())) # Uncomment if you have no write access to R path

repo <- "http://cran.wu.ac.at"
lib.loc <- Sys.getenv("R_LIBS_USER")

update.packages(
    lib.loc, 
    repos = repo,
    ask = FALSE
)

.cran_libs <- c(
  "knitr", # A General-Purpose Package for Dynamic Report Generation in R
  "kableExtra", # Construct Complex Table with 'kable' and Pipe Syntax
  "rmarkdown", # Dynamic Documents for R
  "extrafont", # for extra figure fonts
  "tidyverse", # for dplyr forcats ggplot2 readr tibble
  "grid", # The Grid Graphics Package
  "magrittr", # pipes
  "scales", # Generic plot scaling methods
  "svglite", # for svg files
  "vegan",
  "car", # Companion to Applied Regression
  "rcompanion", #Functions to Support Extension Education Program Evaluation
  "multcomp", # Simultaneous Inference in General Parametric Models 
  "nlme", # Fit Linear Model Using Generalized Least Squares
  "ggResidpanel", # Panels and Interactive Versions of Diagnostic Plots using 
  "lsmeans", # Least-Squares Means
  "survminer", # For Survival Analysis
  "ranger",
  "readr",
  "forcats",
  "dplyr",
  "ggfortify",
  "gtsummary", 
   "cmprsk",  # Competing Risks Packages
  "sm",  # Smooth survival plot - quantile of survival
  "prodlim",
  "pec"
) 

.inst <- .cran_libs %in% installed.packages()
if (any(!.inst)) {
   install.packages(.cran_libs[!.inst],
                    repos = repo,
                    lib = lib.loc)
}

.bioc_libs <- c(
  #"multtest", #Resampling-based multiple hypothesis testing
)

.bioc_inst <- .bioc_libs %in% installed.packages()
if (any(!.bioc_inst)) {
   if (!requireNamespace("BiocManager", quietly = TRUE))
   install.packages("BiocManager")
   BiocManager::install(ask = F, lib = lib.loc)  # upgrade bioC packages
   BiocManager::install(.bioc_libs[!.bioc_inst], ask = F, lib = lib.loc)
}

.local_libs <- c()

.inst <- names(.local_libs) %in% installed.packages()
if (any(!.inst)) {
   install.packages(paste0("~/R/", .local_libs[!.inst]) ,repos = NULL, type = "source", lib = lib.loc)
}

.github_libs <- c()

.github_lib_names <- stringr::str_replace(.github_libs, ".*/(.*)$", "\\1")

.github_inst <- .github_lib_names %in% installed.packages()
if (any(!.github_inst)) {
  devtools::install_github(.github_libs[!.github_inst],
                           lib = lib.loc,
                           dependencies = TRUE)
}

# Load packages into session, and print package version
(loaded.libs <- sapply(c(.cran_libs, .bioc_libs, names(.local_libs), .github_lib_names), require, character.only = TRUE))
if (!all(loaded.libs)) {stop(paste("Package(s):", names(loaded.libs[loaded.libs == FALSE]), "could not be loaded"))}
sapply(c(.cran_libs, .bioc_libs, names(.local_libs), .github_lib_names), packageVersion)
```

```{r style settings, include=F}
options(width = 90, knitr.table.format = "html") 
opts_chunk$set(
  warning = FALSE,
  message = FALSE,
  cache = TRUE,
  dev = "svglite",
  fig.ext = "svg",
  dpi = 300,
#  fig.width = 12,
#  fig.height = 8,
  cache.path = "Survival analysis analysis_loss_cache/",
  fig.path = "Weight_loss_figs/"
)
f_name <- "DejaVu Sans" #sub("\\s//", "", f_name)
f_size <- 14
font_import(pattern = "DejaVuSans", prompt = FALSE)
loadfonts() # registers fonts
theme_set(theme_bw(base_size = f_size, base_family = f_name)) # set theme for plots
pom4 <- ggpomological:::pomological_palette[c(2, 9, 3, 1)] # set colours
```


**Load data for E. pulchripe survival analysis**
```{r load Epibolus survival analysis data, cache = T}

read_csv("3_Epibolus_pul_weight.csv",
         trim_ws = TRUE,
         col_names = TRUE) %>% # read the csv file
  mutate_at(c("Code",
              "Treatment",
              "Replicate",
              "Individual"),
            ~ (factor(.))) %>% # convert to factor
   mutate(Treatment = fct_relevel(Treatment, "Control", "Sterile", "2X Diluted", "Undiluted")) -> # arrange factor levels 4 plotting
  Epibolus_weight
```

# **Fitting survival curves**

*Estimation of the median follow-up time*
```{r Epibolus Estimation of the median,cache=T}
quantile(prodlim(Hist(Time,Status)~ 1, data=Epibolus_weight,reverse=TRUE))
```
No median potential follow-up time of the study. 

*Kaplan-Meier estimates of the probability of survival over time*
```{r Epibolus Kaplan-Meier analysis, cache=T}
km01 <- prodlim(Hist(Time,Status)~1,data=Epibolus_weight)
plot(km01)
 
# To do some fine tuning:

km01 <- prodlim(Hist(Time,Status)~1,data=Epibolus_weight)
par(mar=c(7,7,5,5),                          # margin of figure
    mgp=c(4,1,0))                            # move axis label away from figure
plot(km01,
     xlab="Days",                           # label for x-axis
     ylab="Absolute risk of death",          # label for y-axis
     type="cuminc",                          # increasing risks = 1-survival instead of decreasing survival
     axis1.at=seq(0,2900,365.25),            # time grid for x-axis
     axis1.labels=0:7,                       # time labels for x-axis
     axis2.las=2,                            # rotate labels of y-axis
     atrisk.dist=1,                          # adjust numbers below the figure
     atrisk.labels="Number of \nanimals: ") # labels for numbers below figure

```

*Kaplan-Meier estimates*
```{r Epibolus Kaplan-Meier estimates, cache=T}
# The first thing to do is to use Surv() to build the standard survival object. 
# The variable time records survival time; status indicates whether the patient’s death was observed (status = 1) or that survival time was censored (status = 0).
# Note that a “+” after the time in the print out of km indicates censoring.

## Kaplan Meier Survival Curve
fit <- survfit(Surv(Time, Status) ~1, data = Epibolus_weight)
summary(fit, Time = c(7, 14, 21, 28, 35, 42))
```

*Kaplan Meyer Plot*
```{r Epibolus Kaplan-Meier Plot, cache=T}
plot(fit, xlab="Days", main = 'Kaplan Meyer Plot') # base graphics is always ready.

ggsurvplot(fit)
```

**Stratified Kaplan-Meier**
```{r Epibolus Stratified Kaplan-Meier,cache=T}
km1 <- prodlim(Hist(Time,Status) ~Treatment,data=Epibolus_weight)
plot(km1,
     atrisk.labels=paste(c("Control", "Sterile", "2X Diluted", "Undiluted", ": ")),
     atrisk.title="",
     xlab="Days",  # label for x-axis
     axis1.at=seq(0,42), # time grid for x-axis
     legend.x="bottomleft", # positition of legend
     legend.cex=0.8, # font size of legend
     legend.title="Antibiotic Treatment Groups\n", # 
     logrank=TRUE) # show log-rank p-value
```

*Next, we look at survival curves by treatment.*
```{r Epibolus Kaplan-Meier estimates survival curves by treatment, cache=T}
sfit <- survfit(Surv(Time, Status) ~Treatment, data = Epibolus_weight)
 
ggsurvplot(sfit, data = Epibolus_weight)
```


*Customized survival curves one*
```{r Epibolus Kaplan-Meier estimates Customized survival curves, cache=T}
# Uber customized survival curves
ggsurv <- ggsurvplot(
           sfit,                     # survfit object with calculated statistics.
           data = Epibolus_weight,             # data used to fit survival curves.
           risk.table = TRUE,       # show risk table.
           pval = TRUE,             # show p-value of log-rank test.
           conf.int = F,         # show confidence intervals for 
                                    # point estimates of survival curves.
           palette = pom4,
           xlim = c(0, 45),         # present narrower X axis, but not affect
                                    # survival estimates.
           xlab = "Time in days",   # customize X axis label.
           break.time.by = 7,     # break X axis in time intervals by 500.
           ggtheme = theme_light(), # customize plot and risk table with a theme.
          risk.table.y.text.col = T,# colour risk table text annotations.
          risk.table.height = 0.25, # the height of the risk table
          risk.table.y.text = FALSE,# show bars instead of names in text annotations
                                    # in legend of risk table.
          ncensor.plot = TRUE,      # plot the number of censored subjects at time t
          ncensor.plot.height = 0.25,
          conf.int.style = "step",  # customize style of confidence intervals
          surv.median.line = "hv",  # add the median survival pointer.
          legend.labs = 
            c("Control", "Sterile", "2X Diluted", "Undiluted")    # change legend labels.
        )
ggsurv
```
Although the four curves and three (except Undiluted) appear to overlap in the first 14 days and 35 days respectively, the control animals clearly have a better chance of surviving more than 42 days.
From the summary, We find that the probability of survival in this study after day 14 and day 35 are 95% and 90% respectively.
The associated lower and upper bounds of the 95% confidence interval are also displayed. The p-value shows that the treatment did not have a significant effect on the survival.

**Log-rank test**
```{r Epibolus Log-rank test,cache=T}
survdiff(Surv(Time, Status)~ Treatment,data = Epibolus_weight)
```

**Median survival**
*overall Kaplan-Meier*
```{r Epibolus Median survival analysis,cache=T}
km01 <- prodlim(Hist(Time, Status)~1,data=Epibolus_weight)
quantile(km01)

# Stratified Kaplan-Meier
km1 <- prodlim(Hist(Time, Status)~ Treatment,data=Epibolus_weight)
quantile(km1)

```


# Cox Proportional Hazards Model Analysis
**Cox Proportional Hazards Model**
Next, I’ll fit a Cox Proportional Hazards Model that makes use of all of the covariates in the data set.

```{r Epibolus Cox Proportional Hazards, cache=T}

# Fit Cox Model using Treatment 
cox <- coxph(Surv(Time, Status) ~ Treatment, data = Epibolus_weight)
summary(cox)
cox_fit <- survfit(cox)
# plot(cox_fit, main = "cph model", xlab="Days")
autoplot(cox_fit)

# Let’s create another model where we analyze all the weight losses in the dataset! This shows us how all the variables, when considered together, act to influence survival.

# Fit Cox Model using weight loss or gain after 7 days 
cox <- coxph(Surv(Time, Status) ~ Treatment + `Weight loss7 (g)`, data = Epibolus_weight)
summary(cox)
cox_fit <- survfit(cox)
#plot(cox_fit, main = "cph model", xlab="Days")
autoplot(cox_fit)

# Fit Cox Model using weight loss or gain after 14 days 
cox <- coxph(Surv(Time, Status) ~ Treatment + `Weight loss14 (g)`, data = Epibolus_weight)
summary(cox)
cox_fit <- survfit(cox)
#plot(cox_fit, main = "cph model", xlab="Days")
autoplot(cox_fit)

# Fit Cox Model using weight loss or gain after 21 days 
cox <- coxph(Surv(Time, Status) ~ Treatment + `Weight loss21 (g)`, data = Epibolus_weight)
summary(cox)
cox_fit <- survfit(cox)
#plot(cox_fit, main = "cph model", xlab="Days")
autoplot(cox_fit)

# Fit Cox Model using weight loss or gain after 28 days 
cox <- coxph(Surv(Time, Status) ~ Treatment + `Weight loss28 (g)`, data = Epibolus_weight)
summary(cox)
cox_fit <- survfit(cox)
#plot(cox_fit, main = "cph model", xlab="Days")
autoplot(cox_fit)


# Fit Cox Model using weight loss or gain after 35 days 
cox <- coxph(Surv(Time, Status) ~ Treatment + `Weight loss35 (g)`, data = Epibolus_weight)
summary(cox)

# Fit Cox Model using weight loss or gain after 42 days 
cox <- coxph(Surv(Time, Status) ~ Treatment + `Weight loss42 (g)`, data = Epibolus_weight)
summary(cox)



# Formatting Cox regression results of Treatment using tbl_regression from the gtsummary package
coxph(Surv(Time, Status) ~ Treatment, data = Epibolus_weight) %>% 
  gtsummary::tbl_regression(exp = TRUE) 

# To demonstrate using the survival package, along with ggplot2 and ggfortify, I’ll fit Aalen’s additive regression model for censored data

aa_fit <-aareg(Surv(Time, Status) ~ Treatment, data = Epibolus_weight)
aa_fit

#summary(aa_fit)  # provides a more complete summary of results
autoplot(aa_fit)
```
1. From the output from treatment, we can conclude that treatment did not have statistically significant coefficients.


**Exponential model**
survreg() fits accelerated failure models, not proportional hazards models. The coefficients are logarithms of ratios of survival times, so a positive coefficient means longer survival.
```{r Epibolus Exponential model}
res.exp <- survreg(Surv(Time, Status) ~ Treatment, data = Epibolus_weight, dist = "exponential")
summary(res.exp)
```
Having treatment >1 accelerates the time to event by a factor of exp(Value) = 19.5, 18.7 and 18.8 for sterile, 2X and Und respectively ( 19.5, 18.7 and 18.8 times shorter survival time compared to the baseline survival).


**Comparison of Kaplan-Meier and Cox regression**
```{r Epibolus Comparison of Kaplan-Meier and Cox regression,cache=T}
km.grade <- prodlim(Surv(Time, Status)~Treatment,data=Epibolus_weight)
cox.grade <- coxph(Surv(Time, Status)~Treatment,data=Epibolus_weight,x=TRUE,y=TRUE)
newdata <- data.frame(Treatment=c("Control", "Sterile", "2X Diluted", "Undiluted"))

### first show Kaplan-Meier without confidence limits

plot(km.grade, lty=1, lwd=3,
      col=pom4, confint=FALSE)

## now add survival estimates based on Cox regression 
plotPredictSurvProb(cox.grade, lty=2,
                    col=pom4,
                    add=TRUE, sort(unique(Epibolus_weight$Time)),
                    newdata=newdata)
mtext("Comparison of univariate  Cox regression and stratified Kaplan-Meier")
```


**Plotting individualized survival probabilities**
```{r Epibolus Plotting individualized survival probabilities,cache=T}
cox2 <- coxph(Surv(Time, Status)~Treatment,data=Epibolus_weight,x=TRUE,y=TRUE)
newdata <- data.frame(Treatment=c("Control", "Sterile", "2X Diluted", "Undiluted"))
plotPredictSurvProb(cox2,
                    sort(unique(Epibolus_weight$Time)),
                    newdata=newdata,
                    col=pom4,
                    legend.title="Antibiotic Curing of Epibolus",
                    legend.legend=c("Control", "Sterile", "2X Diluted", "Undiluted"))
mtext("Individualized survival curves from multiple Cox regression
",line=1.5,cex=1.3)
```
