---
title: "Microbiome curing through antibiotics"
subtitle: "Colony counts analysis"
author: "Nweze Julius"
date: "`r Sys.Date()`"
link-citations: yes
csl: fems-microbiology-ecology.csl
output:
  rmarkdown::html_document:
    toc: true
    toc_float: true
    toc_depth: 5
    keep_md: true
    number_sections: false
    highlight: "pygments"
    theme: "flatly"
    dev: "png"
    df_print: "kable"
    fig_caption: true
    code_folding: "show"
editor_options: 
  chunk_output_type: console
---

```{r libraries, include=F}
# Load libraries
#.libPaths(c('~/R/library', .libPaths())) # Uncomment if you have no write access to R path

repo <- "http://cran.wu.ac.at"
lib.loc <- Sys.getenv("R_LIBS_USER")

update.packages(
    lib.loc, 
    repos = repo,
    ask = FALSE
)

.cran_libs <- c(
  "knitr", # A General-Purpose Package for Dynamic Report Generation in R
  "kableExtra", # Construct Complex Table with 'kable' and Pipe Syntax
  "rmarkdown", # Dynamic Documents for R
  "extrafont", # for extra figure fonts
  "tidyverse", # for dplyr forcats ggplot2 readr tibble
  "grid", # The Grid Graphics Package
  "magrittr", # pipes
  "scales", # Generic plot scaling methods
  "svglite", # for svg files
  "vegan",
  "psych",
  "car", # Companion to Applied Regression
  "rcompanion", #Functions to Support Extension Education Program Evaluation
  "multcomp", # Simultaneous Inference in General Parametric Models 
  "nlme", # Fit Linear Model Using Generalized Least Squares
  "ggResidpanel", # Panels and Interactive Versions of Diagnostic Plots using 
  "lsmeans" # Least-Squares Means
) 

.inst <- .cran_libs %in% installed.packages()
if (any(!.inst)) {
   install.packages(.cran_libs[!.inst],
                    repos = repo,
                    lib = lib.loc)
}

.bioc_libs <- c(
  #"multtest", #Resampling-based multiple hypothesis testing
)

.bioc_inst <- .bioc_libs %in% installed.packages()
if (any(!.bioc_inst)) {
   if (!requireNamespace("BiocManager", quietly = TRUE))
   install.packages("BiocManager")
   BiocManager::install(ask = F, lib = lib.loc)  # upgrade bioC packages
   BiocManager::install(.bioc_libs[!.bioc_inst], ask = F, lib = lib.loc)
}

.local_libs <- c()

.inst <- names(.local_libs) %in% installed.packages()
if (any(!.inst)) {
   install.packages(paste0("~/R/", .local_libs[!.inst]) ,repos = NULL, type = "source", lib = lib.loc)
}

.github_libs <- c()

.github_lib_names <- stringr::str_replace(.github_libs, ".*/(.*)$", "\\1")

.github_inst <- .github_lib_names %in% installed.packages()
if (any(!.github_inst)) {
  devtools::install_github(.github_libs[!.github_inst],
                           lib = lib.loc,
                           dependencies = TRUE)
}

# Load packages into session, and print package version
(loaded.libs <- sapply(c(.cran_libs, .bioc_libs, names(.local_libs), .github_lib_names), require, character.only = TRUE))
if (!all(loaded.libs)) {stop(paste("Package(s):", names(loaded.libs[loaded.libs == FALSE]), "could not be loaded"))}
sapply(c(.cran_libs, .bioc_libs, names(.local_libs), .github_lib_names), packageVersion)
```

```{r style settings, include=F}
options(width = 90, knitr.table.format = "html") 
opts_chunk$set(
  warning = FALSE,
  message = FALSE,
  cache = TRUE,
  dev = "svglite",
  fig.ext = "svg",
  dpi = 300,
#  fig.width = 12,
#  fig.height = 8,
  cache.path = "Colony_counts_cache/",
  fig.path = "Colony_counts_figs/"
)
f_name <- "DejaVu Sans" #sub("\\s//", "", f_name)
f_size <- 14
font_import(pattern = "DejaVuSans", prompt = FALSE)
loadfonts() # registers fonts
theme_set(theme_bw(base_size = f_size, base_family = f_name)) # set theme for plots
pom4 <- ggpomological:::pomological_palette[c(2, 9, 3, 1)] # set colours
```

```{r functions, include=F}
PlotLmResid <- function(lm.df, which = c(1:6), mfrow = c(3, 2)){

  if (length(levels(as.factor(lm.df$.fitted))) < 10) {# if number of unique x values is <10 just draw a line through the means
    smoother <- stat_summary(fun.y = mean, colour = "red", geom = "line")
  } else smoother <- stat_smooth(method = "loess", geom = "smooth", se = FALSE, colour = "firebrick", size = 1)
  
  # residuals vs fitted
  g1 <- ggplot(lm.df, aes(.fitted, .resid)) +
    geom_point()  +
    smoother + 
    geom_hline(yintercept = 0, linetype = 2, size = .2) +
    scale_x_continuous("Fitted Values") +
    scale_y_continuous("Residual") +
    labs(title = "Residuals vs Fitted")
  
  # normal qq
  a <- quantile(lm.df$.stdresid, c(0.25, 0.75), na.rm = TRUE)
  b <- qnorm(c(0.25, 0.75))
  slope <- diff(a)/diff(b)
  int <- a[1] - slope * b[1]
  g2 <- ggplot(lm.df, aes(sample = .stdresid)) +
    stat_qq() +
    geom_abline(slope = slope, intercept = int, colour = "firebrick", size = 1) +
      scale_x_continuous("Theoretical Quantiles") +
      scale_y_continuous("Standardized Quantiles") +
      labs(title = "Normal Q-Q")
 
  # scale-location
  g3 <- ggplot(lm.df, aes(.fitted, sqrt(abs(.stdresid)))) +
    geom_point() +
    smoother +
    scale_x_continuous("Fitted Values") +
    scale_y_continuous("Root of Standardized Residuals") +
    labs(title = "Scale-Location")
 
  # residuals vs leverage
  g4 <- ggplot(lm.df, aes(factors, .stdresid)) +
    geom_point() +
    smoother +
    geom_hline(yintercept = 0, linetype = 2, size = .2) +
    scale_x_continuous("Factor Level Combinations") +
    scale_y_continuous("Standardized Residuals") +
    labs(title = "Residuals vs Factor Levels")
 
#   # cook's distance
#   g5 <-  ggplot(lm.df, aes(rows, .cooksd, ymin=0, ymax=.cooksd)) +
#     geom_point() + geom_linerange() +
#     scale_x_continuous("Observation Number") +
#     scale_y_continuous("Cook's distance") +
#     labs(title="Cook's Distance")  
  
  # cooksd vs leverage
  g5 <- ggplot(lm.df, aes(factors, .cooksd)) +
    geom_point() +
    smoother +
    scale_x_continuous("Factor Level Combinations") +
    scale_y_continuous("Cook's distance") +
    labs(title = "Cook's dist vs Leverage")
  
  # g6 <- PlotACF(lm.df)
  bw <- diff(range(lm.df$.resid)) / (2 * IQR(lm.df$.resid) / length(lm.df$.resid) ^ (1/3))
  sshist <- function(x){ # optimise bins
  # 2006 Author Hideaki Shimazaki
  # Department of Physics, Kyoto University
  # shimazaki at ton.scphys.kyoto-u.ac.jp
	N <- 2 : 100
	C <- numeric(length(N))
	D <- C

	for (i in 1:length(N)) {
		D[i] <- diff(range(x)) / N[i]

		edges = seq(min(x), max(x), length=N[i])
		hp <- hist(x, breaks = edges, plot=FALSE)
		ki <- hp$counts

		k <- mean(ki)
		v <- sum((ki-k) ^ 2) / N[i]

		C[i] <- (2 * k-v) / D[i] ^ 2	#Cost Function
	}

	idx <- which.min(C)
	optD <- D[idx]

	bins <- seq(min(x), max(x), length=N[idx])
	# h = hist(x, breaks = bins)
	# rug(x)

	return(bins)
  }
  
  bins <- sshist(lm.df$.resid)
  g6 <- ggplot(lm.df, aes(.resid)) + 
    geom_histogram(breaks = bins)
 
  plots <- list(g1, g2, g3, g4, g5, g6)
 
  # making the plots
  grid.newpage()
 
  if (prod(mfrow) > 1) {
    mypos <- expand.grid(1:mfrow[1], 1:mfrow[2])
    mypos <- mypos[with(mypos, order(Var1)), ]
    pushViewport(viewport(layout = grid.layout(mfrow[1], mfrow[2])))
    formatter <- function(.){}
  } else {
    mypos <- data.frame(matrix(1, length(which), 2))
    pushViewport(viewport(layout = grid.layout(1, 1)))
    formatter <- function(.) {
      .dontcare <- readline("Hit <Return> to see next plot: ")
      grid.newpage()
    }
  }
 
  j <- 1
  for (i in which) {
    formatter()
    print(plots[[i]], vp = viewport(layout.pos.row = mypos[j, ][1], layout.pos.col = mypos[j, ][2]))
    j <- j + 1
  }
}

PlotACF <- function(lm.df){
  ## generate ACF plot for lm and lme
  # compute acf without plotting
  acz <- acf(lm.df$.resid, plot = F)
  # convert to data frame
  acd <- data.frame(lag = acz$lag, acf = acz$acf)
  # use data frame for ggplot
  ggplot(acd, aes(lag, acf)) + 
    geom_bar(colour = "black", fill = "black", stat = "identity", position = "dodge", width = 0.01) +
    geom_point(colour = "black") +
    geom_hline(yintercept = c(0.05, -0.05), linetype = "dashed") +
    geom_hline(yintercept = 0)
}
```
[roey.angel@bc.cas.cz](mailto: roey.angel@bc.cas.cz)  

## Colony counts analysis
Analysis of colony counts analysis over time in response to antibiotics treatment.
Species tested: *G. connexa*


### *G. connexa* colony counts analysis 

**Load data **
```{r load G. connexa data, cache = T}
read_csv("8_Glomeris_colony_counts.csv",
         trim_ws = TRUE,
         col_names = TRUE) %>% # read the csv file
  mutate_at(c("Code",
              "Treatment",
              "Replicate",
              "Individual"),
            ~ (factor(.))) %>% # convert to factor
  mutate_at(c("Date"), ~ as.Date(., "%d/%m/%Y")) %>% # convert to date 
  mutate(Treatment = fct_relevel(Treatment, "Control", "Sterile", "2X Diluted", "5X Diluted")) -> # arrange factor levels 4 plotting
  Glomeris_colony_counts
```
#### Plot time series

**Plot time series of all replicates**
```{r plot all weights 2, cache=TRUE}
ggplot(Glomeris_colony_counts, aes(x = Day, y = `No. of colonies per pellet`, colour = Treatment, group = Code)) + 
  geom_point() +
  geom_line() +
  scale_colour_manual(values = pom4, name = "") +
  scale_y_continuous(label=scientific_10)
```

**Plot time series of means**
```{r plot mean+SE colony counts 2, cache=TRUE}
Glomeris_colony_counts %>%
  filter(`No. of colonies per pellet` > 0) %>%
  drop_na() %>%
  group_by(Treatment, Day) %>%
  summarise(
    `Mean colony counts per pellet` = exp(mean(log(`No. of colonies per pellet`))),
    StD = sd(`No. of colonies per pellet`),
    n = n(),
    SE = StD / sqrt(n)
  ) -> Glomeris_colony_counts_mean

# Create plot
ggplot(Glomeris_colony_counts_mean, aes(x = Day, y = `Mean colony counts per pellet`)) +
  geom_ribbon(
    aes(ymin = `Mean colony counts per pellet` - SE, ymax = `Mean colony counts per pellet` + SE, fill = Treatment),
    alpha = 1 / 4
  ) +
  geom_point(aes(colour = Treatment), size = 3) +
  geom_line(aes(colour = Treatment)) +
  labs(y = "No. of colonies per pellet") +
  scale_colour_manual(values = pom4, name = "") +
  scale_fill_manual(values = pom4, name = "") +
  scale_y_continuous(label = scientific_format(scale = 10))  # Use `scientific_format` from scales package

```

```{r table mean+SE weights 2, cache=TRUE}
Glomeris_colony_counts_mean %>% # output a table of group means
  kable(., digits = c(0, 1, 1, 1, 0, 1)) %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive"), full_width = F)
```
#### Anova tests
**Compare time 0 colony counts between groups**
```{r t0 colony counts aov 2, cache=T}
Glomeris_colony_counts %>% 
  filter(., Day  == 0) %>% # select only day 0 entries 
  aov(`No. of colonies per pellet` ~ Treatment, .) -> # fit an anova model
  aov_t0
  
print(Anova(aov_t0)) # print results

aov_t0_df <- fortify(aov_t0) # make a data frame for plotting
factor.combinations <-
  as.numeric(factor(aov_t0_df$Treatment),
             levels = unique(as.character(paste(aov_t0_df$Treatment)))
  ) # needed for "residuals vs leverage
mod_data_df <-
  cbind(aov_t0_df,
        rows = 1:nrow(aov_t0_df),
        factors = factor.combinations)

PlotLmResid(mod_data_df) # plot residuals
```
The ANOVA model indicates no significant different in the colony counts of the animals' faecal pellets at time 0

**Post-hoc tests for time 0**
```{r t0 colony counts post-hoc 2, cache=T}
TukeyHSD(aov_t0) # Compute Tukey Honest Significant Differences
```
Also in the post-hoc tests shows no signififcant differences.

**Compare time 7 colony counts between groups**
```{r t7 colony counts aov 2, cache=T}
Glomeris_colony_counts %>% 
  filter(., Day  == 7) %>% # select only day 7 entries 
  aov(`No. of colonies per pellet` ~ Treatment, .) -> # fit an anova model
  aov_t7
  
print(Anova(aov_t7)) # print results

aov_t7_df <- fortify(aov_t7) # make a data frame for plotting
factor.combinations <-
  as.numeric(factor(aov_t7_df$Treatment),
             levels = unique(as.character(paste(aov_t7_df$Treatment)))
  ) # needed for "residuals vs leverage
mod_data_df <-
  cbind(aov_t7_df,
        rows = 1:nrow(aov_t7_df),
        factors = factor.combinations)

PlotLmResid(mod_data_df) # plot residuals
```
The ANOVA model indicates a significant difference in the colony counts of the animals' faecal pellets at time 7.

**Post-hoc tests for time 7**
```{r t7 colony counts post-hoc 2, cache=T}
TukeyHSD(aov_t7) # Compute Tukey Honest Significant Differences
```
Pairs: Post-hoc tests show signififcant differences between 2X Diluted-Control and 5X Diluted-Control, but no signififcant differences between Sterile-Control, 2X Diluted-Sterile, 5X Diluted-Sterile and 5X Diluted-2X Diluted.

**Compare time 14 colony counts between groups**
```{r t14 colony counts aov 2, cache=T}
Glomeris_colony_counts %>% 
  filter(., Day  == 14) %>% # select only day 14 entries 
  aov(`No. of colonies per pellet` ~ Treatment, .) -> # fit an anova model
  aov_t14
  
print(Anova(aov_t14)) # print results

aov_t14_df <- fortify(aov_t14) # make a data frame for plotting
factor.combinations <-
  as.numeric(factor(aov_t14_df$Treatment),
             levels = unique(as.character(paste(aov_t14_df$Treatment)))
  ) # needed for "residuals vs leverage
mod_data_df <-
  cbind(aov_t14_df,
        rows = 1:nrow(aov_t14_df),
        factors = factor.combinations)

PlotLmResid(mod_data_df) # plot residuals
```
The ANOVA model indicates no significant differences at day 14.




# Total colony plot
**Plot time series of means for total colony counts**
```{r plot mean+SE colony counts 2, cache=TRUE}
Glomeris_colony_counts %>%
  filter(`Total colonies` > 0) %>%
  drop_na() %>%
  group_by(Treatment, Day) %>%
  summarise(
    `Mean total colonies` = geometric.mean(`Total colonies`),
    StD = sd(`Total colonies`),
    n = n(),
    SE = StD / sqrt(n)
  ) -> Glomeris_total_colony_counts_mean

# Create plot
ggplot(Glomeris_total_colony_counts_mean, aes(x = Day, y = `Mean total colonies`)) +
  geom_ribbon(
    aes(ymin = `Mean total colonies` - SE, ymax = `Mean total colonies` + SE, fill = Treatment),
    alpha = 1 / 4
  ) +
  geom_point(aes(colour = Treatment), size = 3) +
  geom_line(aes(colour = Treatment)) +
  labs(y = "Mean total colonies") +  # Corrected y-axis label
  scale_colour_manual(values = pom4, name = "") +
  scale_fill_manual(values = pom4, name = "") +
  scale_y_continuous(label = scientific_format(scale = 10))  # Use `scientific_format` from scales package

```

```{r table mean+SE weights 2, cache=TRUE}
Glomeris_colony_counts_mean %>% # output a table of group means
  kable(., digits = c(0, 1, 1, 1, 0, 1)) %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive"), full_width = F)
```
#### Anova tests
**Compare time 2 colony counts between groups**
```{r t0 colony counts aov 2, cache=T}
Glomeris_colony_counts %>% 
  filter(., Day  == 2) %>% # select only day 0 entries 
  aov(`Total colonies` ~ Treatment, .) -> # fit an anova model
  aov_t2
  
print(Anova(aov_t2)) # print results

aov_t2_df <- fortify(aov_t2) # make a data frame for plotting
factor.combinations <-
  as.numeric(factor(aov_t2_df$Treatment),
             levels = unique(as.character(paste(aov_t2_df$Treatment)))
  ) # needed for "residuals vs leverage
mod_data_df <-
  cbind(aov_t2_df,
        rows = 1:nrow(aov_t2_df),
        factors = factor.combinations)

PlotLmResid(mod_data_df) # plot residuals
```
The ANOVA model indicates no significant different in the colony counts of the animals' faecal pellets at time 2

**Post-hoc tests for time 0**
```{r t0 colony counts post-hoc 2, cache=T}
TukeyHSD(aov_t2) # Compute Tukey Honest Significant Differences
```
Also in the post-hoc tests shows no signififcant differences.

**Compare time 28 colony counts between groups**
```{r t7 colony counts aov 2, cache=T}
Glomeris_colony_counts %>% 
  filter(., Day  == 28) %>% # select only day 7 entries 
  aov(`Total colonies` ~ Treatment, .) -> # fit an anova model
  aov_t28
  
print(Anova(aov_t28)) # print results

aov_t28_df <- fortify(aov_t28) # make a data frame for plotting
factor.combinations <-
  as.numeric(factor(aov_t28_df$Treatment),
             levels = unique(as.character(paste(aov_t28_df$Treatment)))
  ) # needed for "residuals vs leverage
mod_data_df <-
  cbind(aov_t28_df,
        rows = 1:nrow(aov_t28_df),
        factors = factor.combinations)

PlotLmResid(mod_data_df) # plot residuals
```
The ANOVA model indicates a significant difference in the colony counts of the animals' faecal pellets at time 28.

**Post-hoc tests for time 28**
```{r t28 colony counts post-hoc 2, cache=T}
TukeyHSD(aov_t28) # Compute Tukey Honest Significant Differences
```
Pairs: Post-hoc tests show signififcant differences between 2X Diluted-Control and 5X Diluted-Control, but no signififcant differences between Sterile-Control, 2X Diluted-Sterile, 5X Diluted-Sterile and 5X Diluted-2X Diluted.